---
title: "Data607 Assignment for Week 3"
author: "Leo Yi"
date: "2020-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### #1. Using the 173 majors listed in fivethirtyeight.com’s College Majors [dataset](https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/), provide code that identifies the majors that contain either "DATA" or "STATISTICS"


```{r, eval = T}
majors <- read.csv(url('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv'), stringsAsFactors = F)
str(majors)

majors$Major[grepl('DATA', majors$Major)]
majors$Major[grepl('STATISTICS', majors$Major)]

```

---

##### #2 Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")



```{r, eval = T}
blob <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'

library(stringr)

foods <- str_extract_all(blob, '[a-z]+\\s[a-z]+|[a-z]+')
unlist(foods)
```

---

##### #3 Describe, in words, what these expressions will match:

(.)\1\1

* This would match the string '\1\1' and the character before it, as long as its not a new line \n

"(.)(.)\\\2\\\1"

* This would match anything like a 4 letter palindrome. one character, another character, the same as the 2nd character, the same as the first character, while surrounded by quotes.

(..)\1

* This would match any two characters followed by the string '\1'

"(.).\\\1.\\\1"

* This would match any character, then any other character, then the first character again, then any other character, then the first character again, while surrounded by quotes.

"(.)(.)(.).*\\\3\\\2\\\1"

* This would match any three characters and any (maybe none) characters in between the first 3 characters reversed, while surrounded by quotes.

---

##### #4 Construct regular expressions to match words that:

Start and end with the same character.

* (.)[a-z]*\\\1

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

* ([a-z]{2})[a-z]*\\\1

Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

* [a-z]&ast;([a-z])[a-z]*\\\1[a-z]&ast;\\\1[a-z]&ast;






