---
title: "Data607 Assignment 7"
author: "Leo Yi"
date: "2020-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Working with XML and JSON in R

This assignment requires us to create objects in html, xml, and json and then load them into separate dataframes to compare.

I chose the subject of woodworking, and selected the top 3 books when searched on amazon.

I created the html, xml, and json in text files and saved them and uploaded them to the web.

Lets load some packages and take a look:

#### Packages

```{r, eval = T, error = F, message = F, warning = F}
library(RCurl)
library(XML)
library(jsonlite)
```

#### HTML

```{r}
html_o <- 'https://raw.githubusercontent.com/dataconsumer101/data607/master/books.html'
html_c <- getURLContent(html_o)
html_l <- readHTMLTable(html_c)
html_df <- html_l[[1]]
head(html_df)
```

#### XML

```{r}
xml_o <- 'https://raw.githubusercontent.com/dataconsumer101/data607/master/books.xml'
xml_c <- getURLContent(xml_o)
xml_p <- xmlParse(xml_c)
xml_df <- xmlToDataFrame(xml_p)
head(xml_df)
```

#### JSON

```{r}
json_o <- 'https://raw.githubusercontent.com/dataconsumer101/data607/master/books.json'
json_r <- fromJSON(json_o)
json_df <- json_r[[1]]
head(json_df)
```

#### Differences in Author

```{r}
html_df[1,]$Author
xml_df[1,]$AUTHOR
xml_df[1,]$AUTHOR2
json_df[1,]$Author
```

* For the most part, the data in each of these formats looks the same-- as long as there is only one value per field.
* If we look at the first book with 2 authors, we can see the differences in how the values appear. This is also influenced by how they were entered into each of the objects.
