---
title: "Data607 Assignment 9"
author: "Leo Yi"
date: "2020-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### NYT Movie Reviews

This week's assignment is to access the NYT web API. We'll get data in JSON format and convert it into a dataframe.

Let's load the packages we'll use:

```{r, warnings = F}
library(httr)
library(jsonlite)
library(dplyr)
```

### Authentication and Query Configuration

We'll need to setup some specific information to use in our API call:

```{r}
key <- 'nrs4EBIHbZqiuZK3GFOAwGFO3vMHEfgY'
path = '/svc/movies/v2/reviews/all.json'
query = list(`api-key` = key,
             offset = 0)
```

### Forming the Dataset

The NYT API allows 10 calls per minute and 4,000 requests per day. If we wanted to gather the entire dataset, we could try to time the loops. Perhaps if we get 429 status code, we can add a minute long pause.

For this exercise, we'll only be looking at 10 requests worth of data, with 20 observations each. That should leave us with a 200 row dataset of the latest movies that have been reviewed by the NYT.

```{r}
for (i in 0:10) {
  
  # set the offset parameter based on the loop number
  query[[2]] <- i * 20

  # modify the url to be called
  url <- modify_url('https://api.nytimes.com', 
                    path = path,
                    query = query)
  
  # call the API
  response <- GET(url)  

  # exit the loop if the response is not a success
  if (response$status_code != 200) {
    print(i)
    break
    }
  
  # extract the raw content from the response
  raw_content <- rawToChar(response$content)
  
  # convert the raw content into a json object
  json <- fromJSON(raw_content)
  
  if (i == 0) {
    # flatten the response into a dataframe
    df <- flatten(json$results)
    
  } else {
    # flatten the response into a dataframe
    df_new <- flatten(json$results)
    
    # combine the new dataframe into the initial one
    df <- bind_rows(df, df_new)
    
    # remove the dataset of this current loop
    rm(df_new)
  }
  
}

# preview the dataset
str(df)
```

