---
title: "Data607 Assignment 2"
author: "Leo Yi"
date: "2020-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Movie Survey

This assignment is a survey of 5 individuals. Each individual was asked to rate the following movies on a scale of 1 to 5.
* Joker
* It 2
* Parasite
* Ready or Not
* Avengers Endgame
* Star Wars: Return of the Jedi  

The survey results were stored in a SQL table in pgAdmin4 using postgres. You can see the SQL code [here](https://github.com/dataconsumer101/data607/blob/master/movies.sql).  

The table was created and then exported as a csv, which was uploaded to GitHub as well. We'll grab the raw results from there:

```{r get, eval = T}
raw <- read.csv(url('https://raw.githubusercontent.com/dataconsumer101/data607/master/movies.csv'))
str(raw)
head(raw)
```  

There are *NA* values within the dataset because every person isn't really expected to see every movie.  

Lets rearrange the data using tidyr so its easier to analyze:

```{r, eval = T, results = F, warning = F, error = F, echo = F}
library(tidyr)
library(dplyr)
library(ggplot2)
```

```{r gather, eval = T}
df <- gather(raw, movie, rating, -respondent)
head(df)
```  

Lets looking at the average rating for each movie:

```{r avg, eval = T}
(g <- group_by(df, movie) %>%
  summarize(avg_rating = mean(rating, na.rm = T)) %>%
  arrange(desc(avg_rating)))
```  

I'm also interested in which movies were most popular, so lets count how many people saw them:
```{r count, eval = T}
df$counter <- ifelse(is.na(df$rating),0,1)
(c <- group_by(df, movie) %>%
  summarize(views = sum(counter)) %>%
    arrange(desc(views)))
```

Finally, lets visualize the results so that its a bit easier to digest.  

```{r viz, eval = T}
ggplot(c, aes(x = movie, y = views)) +
  geom_col() +
  ggtitle('Movies by Popularity') +
  theme_classic()  

ggplot(g, aes(x = movie, y = avg_rating)) +
  geom_col() +
  ggtitle('Average Rating of Movie Survey') +
  theme_classic()
```

##### It looks like a survey of 5 people isn't necessarily going to be a great representation of the population.


