---
title: "Data607 Presentation"
author: "Leo Yi"
date: "2020-02-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***
### Introduction
* Business Intelligence Analyst at OnDeck

### Data Reduction

* One way to analyze data
* Summary of a larger dataset
* Easier to digest
* Fewer rows
* See results rather than data points
* Less detailed

***

#### Do You Know Where America Stands On Guns?
* This is a FiveThirtyEight project found [here](https://projects.fivethirtyeight.com/guns-parkland-polling-quiz/)
* The article asks you to guess the level of support for different measures regarding gun control
  + What share of Americans support stricter gun laws?
  + What share of Americans support raising the minimum purchase age to 21?
  + What share of Americans support background checks for all gun purchases?
  + What share of Americans support blocking gun sales to people with a history of mental illness?
  + What share of Americans support arming teachers?
  + What share of Americans support banning assault weapons?
  + What share of Americans support banning high-capacity ammunition magazines?
  + What share of Americans support outlawing all guns?
* The answers are summaries of multiple surveys
  + Overall % of Americans who support the subject
  + Republican support
  + Democratic support
* We'll be looking at the raw data found on GitHub.  

***
#### Using R for Data Reduction
First, we'll load the data into a dataframe object:
```{r load, eval = T}
df <- read.csv(url('https://raw.githubusercontent.com/fivethirtyeight/data/master/poll-quiz-guns/guns-polls.csv'), stringsAsFactors = F)
```  

Next, we'll take a peek at the dataset:
```{r check, eval = T}
str(df)
head(df)
```  

For now, we only want to see the overall, republican, and democratic support levels for each of the questions.
We'll use the dplyr package to reduce the data to look at the average support by question:
```{r lib, eval = T, message = F, warning = F, results = F}
library(dplyr)
```

```{r group, eval = T}
reduced <- group_by(df, Question) %>%
  summarize(overall = mean(Support),
            republican = mean(Republican.Support),
            democratic = (mean(Democratic.Support)))
reduced
```  

We can also include the population, which shows whether the respondents are registered voters:
```{r reg_v, eval = T}
reduced_pop <- group_by(df, Question, Population) %>%
  summarize(overall = mean(Support),
            republican = mean(Republican.Support),
            democratic = (mean(Democratic.Support))) %>%
  arrange(Question, Population)
reduced_pop
```








